### 2023-09-01 sections/2-methods.tex
switch to new figure
still losing it

With these concepts defined, we can now pose the $k$- and $\rho$- nearest neighbors search problems.
Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
the set $S_q$ such that  $|S_q| = k$ and $\forall x \in X \setminus S_q$, $f(q, x) < \max\{f(y, q): y \in S_q \}$; that is,
for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.

@@

\begin{algorithm} % enter the algorithm environment
\caption{Cluster Partition} % give the algorithm a caption
\label{alg:partition} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $cluster$
    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
    \STATE $seeds \leftarrow m$ random points from $cluster.points$
    \STATE $c \leftarrow$ geometric median of $seeds$
    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
    \IF{$|left| > 1$}
        \STATE Partition($left$)
    \ENDIF
    \IF{$|right| > 1$}
        \STATE Partition($right$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/geometry/sketchometry_final.png}
    \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk}
    \label{fig:methods:scaling_behavior}
\end{figure}

@@

\begin{algorithm} 
    \caption{$\rho$-NN(\emph{clusters, query, r})} 
    \label{alg:rnn} 
    \begin{algorithmic}[2]
        \REQUIRE $r \geq 0$
        \REQUIRE $clusters \neq \emptyset$
        \STATE $results \leftarrow \emptyset$
        \IF{$clusters.left \land clusters.left \cap B_X(query, r) \neq \emptyset$}
            \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
                \STATE $\rho$-NN($clusters.left, query, r$)
            \ENDIF
        \ENDIF
        \IF{$clusters.right \land clusters.right \cap B_X(query, r) \neq \emptyset$}
            \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
                \STATE $\rho$-NN($clusters.right, query, r$)
            \ENDIF
        \ENDIF
        \IF{$\neg$$clusters.left$ $\land \neg$$clusters.right$}
            \FOR{$p \in clusters.points$}
                \IF{$p.\delta \leq r$}
                    \STATE $results.push((p, r))$
                \ENDIF
            \ENDFOR
        \ENDIF
        \STATE Return $results$
    \end{algorithmic}
    \end{algorithm}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated$\rho$-NN(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[4] % enter the algorithmic environment
        \STATE $H \Leftarrow$ $\emptyset$
        \STATE $m \Leftarrow$ $\frac{root.radius}{root.cardinality}$
        \WHILE {$|H| = 0$}
            \STATE $H.push(\rho$-NN$(tree, query, m)$)
            \STATE $m \Leftarrow 2m$
        \ENDWHILE
        \WHILE {$|H| < k$}
            \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
            \STATE $\mu \leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
            \STATE $m \leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
            \STATE $H.push(\rho$-NN$(tree, query, m)$)
        \ENDWHILE
        \STATE $H.sort()$
        \STATE Return $H[0], H[1], \cdots H[k-1]$
    \end{algorithmic}
    \end{algorithm}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{QuicksortPartition(\emph{Q, k, l, r})} % give the algorithm a caption
    \label{alg:quicksort-partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[5] % enter the algorithmic environment
        \IF{$l \geq r$}
            \STATE Return $min(l, r)$
        \ELSE 
            \STATE $pivot = l + (r - l) / 2$
            \STATE $Q.swap(pivot, r)$
            \STATE $a \leftarrow l$
            \STATE $b \leftarrow l$
            \WHILE {$b < r$}
                \IF {$Q[b].\delta_{max} < Q[r].\delta_{max}$}
                    \STATE $Q.swap(a, b)$
                    \STATE $a \leftarrow a + 1$
                \ENDIF
                \STATE $b \leftarrow b + 1$
            \ENDWHILE
            \STATE $Q.swap(a, r)$
            \STATE $p \leftarrow r$
            \STATE $g \leftarrow \sum_{i=0}^{p} Q[i].cardinality$
            \IF {$g = k$}
                \STATE Return $p$
            \ELSIF {$g < k$}
                \STATE $QuicksortPartition(Q, k, p + 1, r)$
            \ELSE 
                \IF {$p > 0 \land g > k + Q[p - 1].cardinality$}
                    \STATE $QuicksortPartition(Q, k, l, p - 1)$
                \ELSIF {$p > 0 \land g = k + Q[p - 1].cardinality$}
                    \STATE Return $p - 1$
                \ELSE
                    \STATE Return $p$
                \ENDIF
            \ENDIF
        \ENDIF
    \end{algorithmic}
    \end{algorithm}

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:sieve} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[6] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ [$tree.root$]
        \WHILE{$|Q| > k$}
            \STATE $i \leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
            \STATE $\tau \leftarrow Q[i].\delta_{max}$
            \FOR {$q \in Q$}
                \IF {$q.\delta_{min} > \tau$}
                    \STATE $Q.pop(q)$
                \ENDIF
            \ENDFOR
            \FOR {$q \in Q$}
                \IF {$q.isLeaf \lor |q.cardinality| \leq k$}
                    \STATE $Q.push(q.points)$
                \ELSE
                    \STATE $[l, r] \leftarrow q.children$
                    \STATE $Q.push([l, r])$   
                \ENDIF
                \STATE $Q.pop(q)$
            \ENDFOR 
        \ENDWHILE
        \STATE Return $Q$
    \end{algorithmic}
    \end{algorithm}

@@

\begin{algorithm} 
\caption{GreedySearch(\emph{tree, query, k})} 
\label{alg:greedy_search} 
\begin{algorithmic}[3]
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $H \leftarrow$ priority queue
    \WHILE{$|H| < k \lor H.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$\neg Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.pop().children$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.pop()$
        \STATE $H.push(leaf)$
        \WHILE{$|H| > k$}
            \STATE $H.pop()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $H$
\end{algorithmic}
\end{algorithm}

@@

To establish the asymptotic complexity of $k$-NN search, we adopt 
the terminology used to describe the asymptotic complexity of $\rho$-NN search 
in \cite{yu2015entropy}
and address \emph{coarse search} and \emph{fine search} separately. 
Coarse search refers to the process of identifying clusters
which are likely to contain one of the $k$ nearest neighbors. Fine search refers to the process
of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.

Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 
as is the case in $\rho$-NN search. Still, this distinction is useful in analyzing 
the asymptotic complexity of $k$-NN search. In particular, we can think of the complexity of 
coarse search as an estimate of how many clusters we examine during search, and 
the complexity of fine search as an estimate of how many clusters we must examine all the points 
from during search, even if we do not actually perform these as two distinct phases.

Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
\cite{yu2015entropy}: if call the number of points within a given 
radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
we assume that for any point $p$, query $q$, and radius 
$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}

For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
if this still holds for knn].

To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
number of clusters we must examine all the points from during search. With $\rho$-NN 
search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
as in \cite{yu2015entropy}. With $k$-NN search, however, 
the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
is query-dependent. Thus, for $k$-NN search, we  adjust the definition of 
$F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.

By the triangle inequality, we have that $F \subseteq B(q, r' + 2r_c)$, 
where $B(q, r' + r_c)$ is the ball of radius $r' + r_c$ centered at $q$.
Hence, $|F| \leq |B(q, r' + 2r_c)|$.

Now we have that \begin{align*} \mathbb{E}_q[|B(q, r' + 2r_c)|] &\leq \gamma|B(p, r' + 2r_c)| \\ %because of the assumption that density at different locations on the manifold differ by only a constant factor%
    &\leq \gamma|B(p, r')|\left(\frac{r' + 2r_c}{r'}\right)^d \label{5}\\ %because d = lfd tells us how the number of additional hits reached scales with a change in radius, and our ratio between radii is (r' + 2r_c)/r'%
    &\leq \gamma^2 \mathbb{E}[|B(q, r')|]\left(\frac{r'+2r_c}{r'}\right)^d \\ %using density bound again to go from a statement about an arbitrary point back to a statement about the query%
    &\leq \gamma^2 k\left(\frac{r'+ 2r_c}{r'}\right)^d  \\   %using the fact that for k-nn, the number points in the query ball is k%
\end{align*}

It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
We rely on this fact in our estimation of $r'$.

We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
er's center with half its radius). While, ordinarily we compute local fractal dimension 
by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.

If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}

Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

>>>>

With these concepts defined, we can now pose the $k$- and $\rho$- nearest neighbors search problems.
Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
the set $S_q$ such that  $|S_q| = k$ and $\forall x \in X \setminus S_q$, $f(q, x) \leq \max\{f(y, q): y \in S_q \}$; that is,
for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.

@@

\begin{algorithm} % enter the algorithm environment
\caption{Cluster Partition} % give the algorithm a caption
\label{alg:partition} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $cluster$
    \STATE $m \Leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
    \STATE $seeds \Leftarrow m$ random points from $cluster.points$
    \STATE $c \Leftarrow$ geometric median of $seeds$
    \STATE $l \Leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
    \STATE $r \Leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
    \STATE $left \Leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
    \STATE $right \Leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
    \IF{$|left| > 1$}
        \STATE Partition($left$)
    \ENDIF
    \IF{$|right| > 1$}
        \STATE Partition($right$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/geometry/geometry.pdf}
    \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk}
    \label{fig:methods:scaling_behavior}
\end{figure}

@@

\begin{algorithm} 
    \caption{$\rho$-NN(\emph{clusters, query, r})} 
    \label{alg:rnn} 
    \begin{algorithmic}[2]
        \REQUIRE $r \geq 0$
        \REQUIRE $clusters \neq \emptyset$
        \STATE $results \Leftarrow \emptyset$
        \IF{$clusters.left \land clusters.left \cap B_X(query, r) \neq \emptyset$}
            \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
                \STATE $\rho$-NN($clusters.left, query, r$)
            \ENDIF
        \ENDIF
        \IF{$clusters.right \land clusters.right \cap B_X(query, r) \neq \emptyset$}
            \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
                \STATE $\rho$-NN($clusters.right, query, r$)
            \ENDIF
        \ENDIF
        \IF{$\neg$$clusters.left$ $\land \neg$$clusters.right$}
            \FOR{$p \in clusters.points$}
                \IF{$p.\delta \leq r$}
                    \STATE $results.push((p, r))$
                \ENDIF
            \ENDFOR
        \ENDIF
        \STATE Return $results$
    \end{algorithmic}
    \end{algorithm}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated$\rho$-NN(\emph{root, query, k})} % give the algorithm a caption
    \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[4] % enter the algorithmic environment
        \STATE $H \Leftarrow$ $\emptyset$
        \STATE $m \Leftarrow$ $\frac{root.radius}{|root|}$
        \WHILE {$|H| = 0$}
            \STATE $H.push(\rho$-NN$(root, query, m)$)
            \STATE $m \Leftarrow 2m$
        \ENDWHILE
        \WHILE {$|H| < k$}
            \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
            \STATE $\mu \Leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
            \STATE $m \Leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
            \STATE $H.push(\rho$-NN$(root, query, m)$)
        \ENDWHILE
        \STATE $H.sort()$
        \STATE Return $H[0], H[1], \cdots H[k-1]$
    \end{algorithmic}
    \end{algorithm}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{QuicksortPartition(\emph{Q, k, l, r})} % give the algorithm a caption
    \label{alg:quicksort-partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[5] % enter the algorithmic environment
        \IF{$l \geq r$}
            \STATE Return $min(l, r)$
        \ELSE 
            \STATE $pivot = l + (r - l) / 2$
            \STATE $Q.swap(pivot, r)$
            \STATE $a \Leftarrow l$
            \STATE $b \Leftarrow l$
            \WHILE {$b < r$}
                \IF {$Q[b].\delta_{max} < Q[r].\delta_{max}$}
                    \STATE $Q.swap(a, b)$
                    \STATE $a \Leftarrow a + 1$
                \ENDIF
                \STATE $b \Leftarrow b + 1$
            \ENDWHILE
            \STATE $Q.swap(a, r)$
            \STATE $p \Leftarrow r$
            \STATE $g \Leftarrow \sum_{i=0}^{p} |Q[i]|$
            \IF {$g = k$}
                \STATE Return $p$
            \ELSIF {$g < k$}
                \STATE $QuicksortPartition(Q, k, p + 1, r)$
            \ELSE 
                \IF {$p > 0 \land g > k + |Q[p - 1]|$}
                    \STATE $QuicksortPartition(Q, k, l, p - 1)$
                \ELSIF {$p > 0 \land g = k + |Q[p - 1]|$}
                    \STATE Return $p - 1$
                \ELSE
                    \STATE Return $p$
                \ENDIF
            \ENDIF
        \ENDIF
    \end{algorithmic}
    \end{algorithm}

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:sieve} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[6] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \Leftarrow$ [$tree.root$]
        \WHILE{$|Q| > k$}
            \STATE $i \Leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
            \STATE $\tau \Leftarrow Q[i].\delta_{max}$
            \FOR {$q \in Q$}
                \IF {$q.\delta_{min} > \tau$}
                    \STATE $Q.pop(q)$
                \ENDIF
            \ENDFOR
            \FOR {$q \in Q$}
                \IF {$q.isLeaf \lor |q| \leq k$}
                    \STATE $Q.push(q.points)$
                \ELSE
                    \STATE $[l, r] \Leftarrow q.children$
                    \STATE $Q.push([l, r])$   
                \ENDIF
                \STATE $Q.pop(q)$
            \ENDFOR 
        \ENDWHILE
        \STATE Return $Q$
    \end{algorithmic}
    \end{algorithm}

@@

\begin{algorithm} 
\caption{GreedySearch(\emph{root, query, k})} 
\label{alg:greedy_search} 
\begin{algorithmic}[3]
    \STATE $Q \Leftarrow$ priority queue
    \STATE $Q.push(root)$
    \STATE $H \Leftarrow$ priority queue
    \WHILE{$|H| < k \lor H.pop.\delta > Q.pop.\delta_{min}$}
        \WHILE{$\neg Q.pop.isLeaf$}
            \STATE $[l, r] \Leftarrow Q.pop().children$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \Leftarrow Q.pop()$
        \STATE $H.push(leaf)$
        \WHILE{$|H| > k$}
            \STATE $H.pop()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $H$
\end{algorithmic}
\end{algorithm}

@@

To bound the complexity of $k$-NN search, we group all of the "Sieve" algorithms (Sieve, 
Sieve with Separate Centers, and Greedy Sieve) together, and handle $k$-NN by Repeated $\rho$-NN separately.

We start by considering the complexity of $k$-NN by Repeated $\rho$-NN, as this  
is a natural extension of the complexity bounds for $\rho$-NN search described in 
\cite{ishaq2019clustered}. For this algorithm, we adopt the terminology used in 
\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
Coarse search refers to the process of identifying clusters
which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
Fine search refers to the process
of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.

% Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 
% as is the case in $\rho$-NN search. Still, this distinction is useful in analyzing 
% the asymptotic complexity of $k$-NN search. In particular, we can think of the complexity of 
% coarse search as an estimate of how many clusters we examine during search, and 
% the complexity of fine search as an estimate of how many clusters we must examine all the points 
% from during search, even if we do not actually perform these as two distinct phases.

% Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
% \cite{yu2015entropy}: if call the number of points within a given 
% radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
% are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
% we assume that for any point $p$, query $q$, and radius 
% $\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}

% For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
% if this still holds for knn].

% To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
% number of clusters we must examine all the points from during search. With $\rho$-NN 
% search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
% where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
% as in \cite{yu2015entropy}. With $k$-NN search, however, 
% the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
% is query-dependent. Thus, for $k$-NN search, we  adjust the definition of 
% $F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
% where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.

% By the triangle inequality, we have that $F \subseteq B(q, r' + 2r_c)$, 
% where $B(q, r' + r_c)$ is the ball of radius $r' + r_c$ centered at $q$.
% Hence, $|F| \leq |B(q, r' + 2r_c)|$.

% Now we have that \begin{align*} \mathbb{E}_q[|B(q, r' + 2r_c)|] &\leq \gamma|B(p, r' + 2r_c)| \\ %because of the assumption that density at different locations on the manifold differ by only a constant factor%
%     &\leq \gamma|B(p, r')|\left(\frac{r' + 2r_c}{r'}\right)^d \label{5}\\ %because d = lfd tells us how the number of additional hits reached scales with a change in radius, and our ratio between radii is (r' + 2r_c)/r'%
%     &\leq \gamma^2 \mathbb{E}[|B(q, r')|]\left(\frac{r'+2r_c}{r'}\right)^d \\ %using density bound again to go from a statement about an arbitrary point back to a statement about the query%
%     &\leq \gamma^2 k\left(\frac{r'+ 2r_c}{r'}\right)^d  \\   %using the fact that for k-nn, the number points in the query ball is k%
% \end{align*}

% It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
% points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
% average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
% We rely on this fact in our estimation of $r'$.

% We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
% $d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
% and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
% er's center with half its radius). While, ordinarily we compute local fractal dimension 
% by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
% a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.

% If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
% \begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}

% Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
% \begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

### 2023-09-01 sections/3-datasets.tex
still losing it

\section{Datasets and Distance Functions}
\label{sec:datasets-and-distance-functions}

Provide details on the datasets we use for benchmarks.

Add interesting t-SNE or UMAP embeddings of datasets.
For example~\ref{fig:discussion:umap-annthyroid-euclidean}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
    \label{fig:discussion:umap-annthyroid-euclidean}
\end{figure}

@@

Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history. This dataset 
is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure [TODO: ADD FIGURE NUMBER] in Section~\ref{section:methods},
this dataset is a good candidate for CAKES because it exhibits low fractal dimension.

@@

deep1b: cosine

fashion-mnist: euclidean

gist: euclidean

glove: cosine

kosarak: jaccard

mnist: euclidean

nytimes: cosine

sift: euclidean

last.fm: cosine

>>>>

\section{Datasets}
\label{sec:datasets-and-distance-functions}

@@

Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history. This dataset 
is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure 1 in Section 2,
this dataset is a good candidate for CAKES because it exhibits low fractal dimension.

@@

We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
which we list below, along with the distance functions with which they are typically used.

\begin{itemize} 
    \item DEEP1B (Cosine): ~1 billion 96-dimensional deep descriptors to 
    \item Fashion-MNIST (Euclidean): 60,000 784-dimensional vector representations of images of clothing items
    \item GIST (Euclidean): 1 million 960-dimensional image descriptors 
    \item GloVe (Cosine): ~1.2 million 25-dimensional vector representations of words
    \item Kosarak (Jaccard): ~75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
    \item MNIST (Euclidean): 60,000 784-dimensional vector representations of images of handwritten digits
    \item NYTimes (Cosine): 290,000 256-dimensional vector representations of words
    \item SIFT (Euclidean): 1 million 128-dimensional image descriptors
    \item Last.fm (Cosine): ~292, 000 65-dimensional vector representations of songs
\end{itemize}

### 2023-09-01 sections/4-results.tex
still losing it

Write about results \dots

@@

Provide link to source code on GitHub.

>>>>

Figure [ADD FIGURE NUMBER] is a violin plot which shows the performance of our algorithms
on [ADD DATASET] under [ADD DISTANCE FUNCTION] distance, using [ADD NUMBER OF SHARDS] shards.
On the horizontal axis, we have each of the $k$-NN algorithms described in Section 2.4, as well as  
an additional "algorithm" which we call "Oracle $\rho$-NN" [REPLACE WITH ACTUAL NAME].  With "Oracle $\rho$-NN", we perform $\rho$-NN search using the 
distance to the $k$-th nearest neighbor as the radius. Of course, as it is impossible to
to know the distance to the $k$-th nearest neighbor \emph{a priori}, this algorithm 
is useful only for benchmarking purposes.

The vertical axis for Figure [ADD FIGURE NUMBER] is the time query in seconds on a logarithmic scale. 
As shown in the key, different colors represent different values of $k$. 
The plot shows the distribution of query times for each algorithm for each value of $k$. 
Because of the logarithmic scale of the vertical axis, it is unsurprising that we 
see longer lower tails than upper tails for many of these algorithms.

@@

https://github.com/URI-ABD/clam
